Memory-Efficient Mixed-Precision Implementations\\for Robust Explicit Model Predictive Control

### 2019-04-12 emsoft-2019/Motiv.tex
new inverted pendulum
Comments in Motiv

\begin{figure}[h]
	\begin{tabular}{cc}
	\includegraphics[width=2cm,height=3cm]{Figs/inv_pend.png}&	\includegraphics[width=6cm,height=4cm]{Figs/regs.jpg}\\
	(a)&(b)
	\end{tabular}
	\caption{(a) Inverted pendulum; (b) 2-D plot of polyhedral partitions for an explicit MPC applied to the inverted pendulum.}
	\label{fig:inverted_pendulum}
\end{figure}

@@

Our overall goal is to 
(i) design a robust MPC controller that achieves the performance objectives including hard constraints in spite of an additional bounded
disturbance $\Delta$ modeling the implementation error (both errors due to choosing a wrong region and the finite-precision computation of the control action);
and at the same time
%\eva{what about the modelling error that $\Delta$ bounded in the paragraph before?}
(ii) to minimize the total number of bits that are required for storing the explicit controller.

@@

For our example, we choose $\Delta_0=0.01$ and $\mathcal W = \{0\}$. 
We use MATLAB to find a robust explicit MPC with disturbance set $\mathcal W_{\Delta_0}$. Over the computed controller, we compute the error due to an incorrect region selection $e$ and in case that $e<\Delta_0$, we ask Daisy
to provide a mixed-precision scheme such that the associated implementation error is at most $\Delta_0$, that is,
at all points in the control domain, the difference between the infinite-precision controller function
and its mixed-precision implementation is at most $\Delta_0$.

Unfortunately, $\Delta_0=0.01$ is not realizable, as the point location error is $0.019$ and hence $e>\Delta_0$. 
%always greater than zero either due to selecting the wrong region in the affine controller function or due to the finite-precision computation of the control action.
%\eva{Is this the error due to mixed-precision or due to wrong regions?}
Therefore, we define $\Delta_1 := \Delta_0 + \varepsilon_{SAFE}$ with $\varepsilon_{SAFE}$ being a fixed threshold (we chose $\varepsilon_{SAFE} = 0.04$) and enlarge the disturbance set as $\mathcal W_{\Delta_1}$. 
Solving the robust explicit MPC problem for this enlarged disturbance set gives the new controller,
shown in Figure~\ref{fig:inverted_pendulum}(b).
The controller has $14$ regions. Following demostrates one sample mapping, including the activation condition and its corresponding affine function
\begin{equation*}
\resizebox{.9\hsize}{!}{$
\begin{bmatrix}
-0.005  &0.999\\
-0.998  &-0.049\\
0.005  &-0.999\\
0.998   &0.049
\end{bmatrix}x_k\leq 
\begin{bmatrix}
0.416\\
3.157\\
0.617\\
-0.0124
\end{bmatrix}
\Rightarrow
u(x_k)=\begin{bmatrix}
0.05&-9.67
\end{bmatrix}x_k+4.02$}
\end{equation*}
where, output of robust EMPC at state $x_k$ is denoted by $u(x_k)$. Notice that all the rest of mappings have similar format.
%\RM{give the control function. I think it will be good to explicitly write the polyhedra and the affine maps}

>>>>

\begin{figure}[h]
	\begin{tabular}{cc}
	\includegraphics[width=2cm,height=3cm]{Figs/inverted_pendulum.pdf}&	\includegraphics[width=6cm,height=4cm]{Figs/regs.jpg}\\
	(a)&(b)
	\end{tabular}
	\caption{(a) Inverted pendulum; (b) 2-D plot of polyhedral partitions for an explicit MPC applied to the inverted pendulum.}
	\label{fig:inverted_pendulum}
\end{figure}

@@

Our overall goal is to 
(i) design a robust MPC controller that achieves the performance objectives including hard constraints in spite of an additional bounded
disturbance $\Delta$ modeling the implementation error (both errors due to choosing a wrong region and the finite-precision computation of the control action);
and at the same time
%\eva{what about the modelling error that $\Delta$ bounded in the paragraph before?}
(ii) to minimize the total number of bits that are required to represent an explicit controller.

@@

For our example, we choose $\Delta_0=0.01$ and $\mathcal W = \{0\}$. 
We use MATLAB to find a robust explicit MPC with disturbance set $\mathcal W_{\Delta_0}$. 
\RM{the following setence is not understandable. This is the first time we introduce incorrect region selection.
Who is Daisy?}
\RM{A better version is to informally define incorrect region selection. Say that we run an algorithm for incorrect region
selection to bound the possible error. 
Break the sentence into two. Introduce Daisy as a mixed-precision tuning tool.}
Over the computed controller, we compute the error due to an incorrect region selection $e$ and in case that $e<\Delta_0$, we ask Daisy
to provide a mixed-precision scheme such that the associated implementation error is at most $\Delta_0$, that is,
at all points in the control domain, the difference between the infinite-precision controller function
and its mixed-precision implementation is at most $\Delta_0$.

Unfortunately, $\Delta_0=0.01$ is not realizable, as the region selection error is $0.019$ and hence $e>\Delta_0$. 
%always greater than zero either due to selecting the wrong region in the affine controller function or due to the finite-precision computation of the control action.
%\eva{Is this the error due to mixed-precision or due to wrong regions?}
Therefore, we define $\Delta_1 := \Delta_0 + \varepsilon_{SAFE}$ with $\varepsilon_{SAFE}$ being a fixed threshold (we chose $\varepsilon_{SAFE} = 0.04$) and enlarge the disturbance set as $\mathcal W_{\Delta_1}$. 
Solving the robust explicit MPC problem for this enlarged disturbance set gives the new controller,
shown in Figure~\ref{fig:inverted_pendulum}(b).
The controller has $14$ regions. Following demostrates one sample mapping, including the activation condition and its corresponding affine function
\begin{equation*}
\resizebox{.9\hsize}{!}{$
\begin{bmatrix}
-0.005  &0.999\\
-0.998  &-0.049\\
0.005  &-0.999\\
0.998   &0.049
\end{bmatrix}x_k\leq 
\begin{bmatrix}
0.416\\
3.157\\
0.617\\
-0.0124
\end{bmatrix}
\Rightarrow
u(x_k)=\begin{bmatrix}
0.05&-9.67
\end{bmatrix}x_k+4.02$}
\end{equation*}
where, output of robust EMPC at state $x_k$ is denoted by $u(x_k)$. Notice that all the rest of mappings have similar format.
%\RM{give the control function. I think it will be good to explicitly write the polyhedra and the affine maps}

### 2019-04-12 emsoft-2019/intro.tex
Comments in Motiv

% !TEX root = main.tex
\section{Introduction}
%\eva{General comment: should we add some more references to related work (and not only to MPC)?}
Model predictive control (MPC) is a technique to design control actions by solving finite-horizon open-loop
optimal control problems at each sampling instant.
The result of each optimization gives a sequence of optimal control actions, only the first of which is applied
to the process.
The same procedure is applied in the next time instant with a shifted time horizon and a new initial state, 
after receiving the updated values of the process state.
The optimization problem in MPC uses a dynamic model of the process, encodes all input and output (state) constraints, and 
optimizes a performance index. 
MPC has shown to be successful in a wide variety of industrial applications, due to its 
ability to systematically handle processes with many state and input variables as well as constraints on them. 
\todo{any citations?}

@@

However, there are still bottlenecks in implementing explicit MPC on resource-constrained embedded micro-controllers.
First, implementations of explicit MPCs can suffer from large memory usage, because the solution of the optimization
problem can involve many (often hundreds) of regions~\cite{memoryMPC}.
Since the memory consumption grows linearly with the number of regions, 
this can be a limiting factor in using explicit MPC on resource-constrained micro-controllers.
Second, many low-end micro-controllers only support fixed-point arithmetic.
Errors in the controller implementation are inversely proportional to the number of bits used for representing each variable~\cite{Anta10}. 
An implementation of explicit MPC has to be \emph{robust} to 
implementation errors to be able to enforce hard constraints on the states at run time~\cite{imperialrmpc}.

@@

Our proposed method iteratively solves a robust version of MPC which explicitly considers finite-precision implementation errors
in the model; see~\autoref{fig:overview}.
Initially, we estimate a bound $\Delta = \Delta_0$ on the implementation error and solve a
min-max quadratic program for explicit robust MPC \cite{delaPea:2005}, where the system
model has an explicit disturbance bounded by $\Delta$.
The solution is represented as a piecewise affine map.
We use an automated mixed-precision tuning tool to find an approximate implementation of the piecewise affine
map. 
We statically compute the error of the implementation, taking into account errors
in evaluating the affine maps as well as choosing the wrong affine map due to quantization.
If the error is at most $\Delta$, we are done: the implementation satisfies the robustness margin in the model.
If not, we increase the error bound $\Delta$ and run the loop again to find a new robust controller and find its best mixed-precision
implementation.

>>>>

% !TEX root = main.tex
\section{Introduction}
%\eva{General comment: should we add some more references to related work (and not only to MPC)?}

Model predictive control (MPC) is a technique to design control actions by solving finite-horizon open-loop
optimal control problems at each sampling instant \cite{RMD17}.
The result of each optimization gives a sequence of optimal control actions, only the first of which is applied
to the process.
The same procedure is applied in the next time instant with a shifted time horizon and a new initial state, 
after receiving the updated values of the process state.
The optimization problem in MPC uses a dynamic model of the process, encodes all input and output (state) constraints, and 
optimizes a performance index. 
MPC has shown to be successful in a wide variety of industrial applications \cite{QinBadgwell03}, due to its 
ability to systematically handle processes with many state and input variables as well as constraints on them.

@@

However, there are still bottlenecks in implementing explicit MPC on resource-constrained embedded micro-controllers.
First, implementations of explicit MPCs can suffer from large memory usage, because the solution of the optimization
problem can involve many (often hundreds) of regions.
Since the memory consumption grows linearly with the number of regions, 
this can be a limiting factor in using explicit MPC on resource-constrained micro-controllers.
Second, many low-end micro-controllers only support fixed-point arithmetic.
Errors in the controller implementation are inversely proportional to the number of bits used for representing each variable~\cite{Anta10}. 
An implementation of explicit MPC has to be \emph{robust} to 
implementation errors to be able to enforce hard constraints on the states at run time~\cite{imperialrmpc}.

@@

Our proposed method iteratively solves a robust version of MPC 
which explicitly considers finite-precision implementation errors as disturbances
in the dynamics \cite{RamirezC06}; see~\autoref{fig:overview}.
Initially, we estimate a bound $\Delta = \Delta_0$ on the implementation error and solve a
min-max quadratic program for explicit robust MPC \cite{delaPea:2005,RamirezC06}, where the system
model has an explicit disturbance bounded by $\Delta$.
The solution is represented as a piecewise affine map.
\RM{need to add the step on incorrect region selection.}
We use an automated mixed-precision tuning tool to find an approximate implementation of the piecewise affine
map. 
We statically compute the error of the implementation, taking into account errors
in evaluating the affine maps as well as choosing the wrong affine map due to quantization.
If the error is at most $\Delta$, we are done: the implementation satisfies the robustness margin in the model.
If not, we increase the error bound $\Delta$ and run the loop again to find a new robust controller and find its best mixed-precision
implementation.
